---
title: "Untitled"
output:
  pdf_document: default
  html_document: default
---

1.descriptive statistics
1.1
```{r,echo=F}
library(dplyr)
library(car)
df<-Salaries
nrow(df)
colnames(df)
```

1.2 qualitative variables
```{r,echo=F}
rank_sum <- summary(df$rank)
discpl_sum <- summary(df$discipline)
sex_sum <- summary(df$sex)
rank_sum
discpl_sum
sex_sum
```

1.3 quantitative variable
1.3.1 salary
```{r,echo=F}
library(ggplot2)
salary_hist <- df %>% ggplot() +
geom_histogram(aes(x = salary), color = "black", fill = "blue", bins = 50) +
labs(x = "Nine-Month Salary (in dollars)", y = "Count", title = "Nine-Month Salary Distribution")

salary_hist
```
```{r,echo=F}
summary(df$salary)
```


1.3.2 years since Ph.D

```{r,echo=F}
phd_hist <- df %>% ggplot() +
geom_histogram(aes(x = yrs.since.phd), color = "black", fill = "green", bins = 50) +
labs(x = "Years since Ph.D.", y = "Count", title = "Years since Ph.D. Distribution")

phd_hist
```

```{r,echo=F}
summary(df$yrs.since.phd)
```

1.3.3 years of service

```{r,echo=F}
service_hist <- df %>% ggplot() +
geom_histogram(aes(x = yrs.service), color = "black", fill = "orange", bins = 50) +labs(x = "Years of Service", y = "Count",title = "Years of Service Distribution")

service_hist
```

```{r,echo=F}
summary(df$yrs.service)
```

2 Assess the relationship between the quantitative variable and qualitative variable

2.1 Assessing years since Ph.D,salary,and rank

```{r,echo=F}
library(ggplot2)
attach(Salaries)
ggplot(mapping=aes(x=yrs.since.phd,y=salary)) +
geom_point(mapping=aes(colour=rank)) 
```

2.2 Assessing years since Ph.D, salary and discipline
```{r,echo=F}
library(ggplot2)
ggplot(mapping=aes(x=yrs.since.phd,y=salary)) +
geom_point(mapping=aes(colour=discipline)) 
```

2.3 Assessing Salary, years of service and rank

```{r,echo=F}
library(ggplot2)
ggplot(mapping=aes(x=yrs.service,y=salary)) +
geom_point(mapping=aes(colour=rank)) 
```

2.4 Assessing salary, sex, years of service
```{r,echo=F}
library(ggplot2)
ggplot(mapping=aes(x=yrs.service,y=salary)) +
geom_point(mapping=aes(colour=sex)) 
```

3. Univariate regression analysis
3.1 check outlier
```{r,echo=F}
boxplot(salary)
boxplot(yrs.service)
boxplot(yrs.since.phd)
```
from the boxplot, we can find that there is no obvious outlier in salary, years of service and years since Ph.D

3.2 salary~rank
```{r,echo=F}
mod_rank_slr <- lm(salary~rank, data = df)
summary(mod_rank_slr)
```

3.3 salary~sex
```{r,echo=F}
mod_sex_slr <- lm(salary~sex, data =df)
summary(mod_sex_slr)
```

3.4 salary~discipline
```{r,echo=F}
mod_discpl_slr <- lm(salary~discipline, data = df)
summary(mod_discpl_slr)
```

3.5 salary~years since Ph.D
```{r,echo=F}
mod_phd_slr <- lm(salary~yrs.since.phd, data = df)
summary(mod_phd_slr)
```

3.6 salary~years of service
```{r,echo=F}
mod_service_slr <- lm(salary~yrs.service,data=df)
summary(mod_service_slr)
```

4 Multivariate regression analysis
4.1 full model
```{r,echo=F}
full_mod <- lm(salary~., data = df)

summary(full_mod)
```

4.2 model without sex
```{r,echo=F}
mod_no_sex <- lm(salary~rank+discipline+yrs.since.phd+yrs.service, data=df)
summary(mod_no_sex)
```

VIF
```{r,echo=F}
car::vif(mod = mod_no_sex)
```

4.3 model without sex, years since Ph.D, years of service
```{r,echo=F}
mod_no_phd <- lm(salary~discipline+yrs.service+rank,
                  data = df)

summary(mod_no_phd)
```

VIF
```{r,echo=F}
car::vif(mod_no_phd)
```

4.4 model without rank and sex

```{r,echo=F}
mod_no_rank <- lm(salary~discipline+yrs.since.phd+yrs.service,
                  data = df)

summary(mod_no_rank)
```

VIF
```{r,echo=F}
car::vif(mod_no_rank)
```

4.5 model without sex and years of service
```{r,echo=F}
mod_no_service <-lm(salary~discipline+yrs.since.phd+rank,
                  data = df)

summary(mod_no_service)
```

VIF
```{r,echo=F}
car::vif(mod_no_service)
```

4.6 final model
```{r,echo=F}
mod_final <- lm(salary~discipline+rank,data=df)
summary(mod_final)
```


5. model diagnostic
```{r,echo=F}
attach(Salaries)
reg1<-lm(salary~rank+discipline)
summary(reg1)
```

```{r,echo=F}
plot(reg1)
```




```{r,echo=F}
salary1<-log(salary)
reg2<-lm(salary1~rank+discipline)
summary(reg2)
plot(reg2)
```


```{r,echo=F}
salary2<-sqrt(salary)
reg3<-lm(salary2~rank+discipline)
summary(reg3)
plot(reg3)
```

```{r,echo=F}
salary3<-1/salary
reg4<-lm(salary3~rank+discipline)
summary(reg4)
plot(reg4)
```

```{r,echo=F}
cooksd=cooks.distance(reg4)
```

```{r,echo=F}
n<-length(salary)
p<-4
m<-4/(n-p)
t<-which(cooksd>m)
t
{plot(cooksd)
abline(h =m, col="red")}  

```

```{r,echo=F}
Salaries1<-Salaries[-t,]
```

```{r,echo=F}
salary4<-1/Salaries1$salary
reg5<-lm(salary4~Salaries1$rank+Salaries1$discipline)
summary(reg5)
plot(reg5)
```

7. estimated factor level means

the mean salary of rank is Assistant professor and discipline is A is 10^5/1.314=76103.5
the mean salary of rank is Assistant professor and discipline is B is 10^5/1.1967=83563.13
the mean salary of rank is Associate professor and discipline is A is 10^5/1.1446=87366.77
the mean salary of rank is Associate professor and discipline is B is 10^5/1.0273=97342.55
the mean salary of rank is professor and discipline is A is 10^5/0.8766=114077.1
the mean salary of rank is professor and discipline is B is 10^5/0.7593=131700.3



```{r,echo=F}
ggplot(data=Salaries1)+
  geom_boxplot(mapping=aes(x=rank,y=salary))+
  facet_wrap(~discipline)
```



